Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-02-04 main.tex
Notes
Formulation of stack safety based on the positive/negative functor. Hyperhyperproperties?

We assume we have an underlying step function that takes machine
states to machine states, $\MACHS \rightharpoonup \MACHS$; we also
assume the only possible partiality is because of a $\HALT$
instruction.

%The
%{\em underlying} step function, $\mathord{\stepstounderfull} \in
%\MACHS \rightharpoonup \MACHS$, embodies the standard RISC-V ISA
%semantics.  \bcp{It's probably clearer not to put RISCV in the
%  notation---for most of the paper, the fact that it's a RISCV really
%  doesn't matter.  Moreover, I think we want to make it a partial
%  function, for simplicity of definitions in this part, and if we
%  insist on RISCV here this will confuse people that know it doesn't
%  have a HALT instruction.}

% \bcp{Needed?}A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
% sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
% $\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.

@@

\iftext
Intuitively, we use the word {\em policy} for any kind of dynamic
enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
capability hardware, etc.). We use {\em property} for a mathematical
predicate on programs (i.e., machine states) articulating some useful
security goal, e.g. noninterference. \iftext\bcp{Not quite true, if we stick
  with hyperproperties.} \fi
\fi

@@

@@

%
%\subsection{Observations}
%
%A {\em component observation} is either a word or the special value $\bot$.
%A {\em state observation} is a map from machine components to component
%observations.
%
%An {\em observation} is a nonempty, finite or infinite sequence of state
%observations in which (1) adjacent elements are always distinct (``no
%stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
%the the $j^{th}$ state observation in the sequence must also be mapped to
%$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
%component's value changes from $\bot$ to anything else, it never changes
%back to $\bot$ again).
%
%In particular, the observation $\obstrace(\seq{(\contour,\mach)})$ arising
%from a contoured trace $\seq{(\contour,\mach)}$ is sequence of state
%observations $\stateobs_0,\, \stateobs_1,\, \ldots$ defined informally as
%follows:
%\begin{itemize}
%\item $\stateobs_0$ maps every visible component (where $\contour_0 = \VIS$)
%to its value in $\mach_0$ and every hidden
%component to $\bot$;
%\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for each
%visible component (where $\contour_{i+1} = \VIS$), copying its
%value from $\mach_{i+1}$, and (2) for each
%hidden component, where $\contour_i = \VIS$, copying its
%value from $\stateobs_{i+1}$\apt{$\stateobs_i$ ?}, and (3) if the resulting $\stateobs_{i+1}$ is
%identical to $\stateobs_{i}$, then skipping $\stateobs_{i+1}$.
%\iftext\bcp{Too informal!!}\fi
%\end{itemize}

@@

%\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

% A {\em contoured (hyper)property} is a function that takes a contour to a
% hyperproperty.  That is, it takes a starting contour and produces a
% \bcp{prefix-closed?} set of pairs of sequences of machine states.

% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
% \times X) $
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a function that allows ``reading off'' a contour from an $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup X$.
% \end{enumerate*}

A {\em trace} of length $n$, where $n$ can be finite or infinite\bcp{awkward}, is a
length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
machine states in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

A {\em contoured hyperproperty} is a set of triples, each consisting of a
contour along with a pair of traces.

\leo{No longer needed:?}
A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
such that each trace in every pair of traces in the set executes the
same instruction sequence.

Finally, we can parameterize the traces by some initial application-specific
state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
trace comes with more information, for example call sites in stack safety.
\bcp{If we use this later, we need to say it more formally to be clear about
  what we mean---e.g., I had thought both the traces and the contour would
  be functions of $X$!}

>>>>

The {\em underlying} step function, $\mathord{\stepstounderfull} \in
\MACHS \rightarrow \MACHS$, embodies the standard RISC-V ISA
semantics.
%\bcp{It's probably clearer not to put RISCV in the
%  notation---for most of the paper, the fact that it's a RISCV really
%  doesn't matter.  Moreover, I think we want to make it a partial
%  function, for simplicity of definitions in this part, and if we
%  insist on RISCV here this will confuse people that know it doesn't
%  have a HALT instruction.}

A {\em trace} is a finite or infinite sequence of machine states,
$\seq{\mach} = \mach_0,\, \mach_1,\ldots$, in which adjacent machine
states are related by $\stepstounderfull$---i.e., in which
$\mach_{i-1} \stepstounderfull \mach_i$ for each $i > 0$ in the
sequence.

@@

\iftext
Intuitively, we use the word {\em policy} for any kind of dynamic
enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
capability hardware, etc.).
\leo{FIX: terminology}
We use {\em property} for a mathematical
predicate on programs (i.e., machine states) articulating some useful
security goal, e.g. noninterference. \iftext\bcp{Not quite true, if we stick
  with hyperproperties.} \fi
\fi

@@

We denote this (reflexive) flows-to relation with $\flows$. Then
$\LOWSEC \flowsstrict \HIGHSEC$ and $\HIGHINT \flowsstrict
\LOWINT$. We can easily lift this relation to pairs of secrecy and
integrity labels, and from there to contours pointwise.

@@

\subsection{Observations}

A {\em component observation} is either a word or the special value $\bot$.
A {\em state observation} is a map from machine components to component
observations.

An {\em observation} is a nonempty, finite or infinite sequence of state
observations in which (1) adjacent elements are always distinct (``no
stuttering'') and (2) any component $\component$ that is mapped to $\bot$ by
the the $j^{th}$ state observation in the sequence must also be mapped to
$\bot$ by all observations earlier than $j$ in the sequence (i.e., once a
component's value changes from $\bot$ to anything else, it never changes
back to $\bot$ again).

In particular, the observation $\obstrace(\contour, \seq{\mach})$
arising from a pair of a contour $\contour$ and a trace $\seq{\mach}$
is sequence of state observations $\stateobs_0,\, \stateobs_1,\,
\ldots$ defined informally as follows:
\begin{itemize}
\item $\stateobs_0$ maps every visible component $\component$ (where
  $\contour(\component) = (\LOWSEC, \_)$) to its value in $\mach_0$
  and every hidden component to $\bot$;
\item $\stateobs_{i+1}$ is defined from $\stateobs_{i}$ by (1) for
  each visible component $\component$ (where $\contour(\component) =
  (\LOWSEC, \_)$), copying its value from $\mach_{i+1}$, and (2) for
  each hidden component $\component$, where $\contour (\component) =
  (\HIGHSEC,\_)$, copying its value from $\stateobs_{i}$, and (3) if
  the resulting $\stateobs_{i+1}$ is identical to $\stateobs_{i}$,
  then skipping $\stateobs_{i+1}$.  \iftext\bcp{Too informal!!}\fi
\end{itemize}

@@

\iftext\bcp{Emphasize that ``observability is not observable.''}\fi

@@

% A {\em contoured (hyper)property} is a function that takes a contour to a
% hyperproperty.  That is, it takes a starting contour and produces a
% \bcp{prefix-closed?} set of pairs of sequences of machine states.
% 
% An {\em observer} $\observer \in CONTOURS \rightarrow (\MA \exists X.\; (\CONTOURS \rightarrow X) \times
% (X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
% \times X) $
% comprises
% \begin{enumerate*}
%   \item a private type $X$,
%   \item a function taking a contour to an initial value of $X$, and
%   \item a function that allows ``reading off'' a contour from an $X$, and
%   \item a partial step function of type
%   $(\MACHS \times X) \rightharpoonup X$.
% \end{enumerate*}

Let $(\stepplus) :: (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS
\times \MACHS)$ be an extension of the underlying step function, such
that $(C,M) \stepplus (C',M')$ implies that $M \stepstounderfull M'$.
Given a contour $C$ and a machine state $M$, the {\em contoured trace}
induced by $C$ and $M$, written $\induced{C}{M}$ is the largest
(finite or infinite) sequence of pairs of (contour, machine state)
pairs, in which adjacent pairs are related by $\stepstounderfull$ and
for which $C \flows C_i$ for all $i > 0$ in the sequence.
\leo{That is, the induced contoured trace by a (C,M) pair incorporates
  the ``return'' snipping}

Let $\initmach$ be the set of initial states of the machine and let
$\initcontour(M)$ be the initial contour assigned to each $M \in
\initmach$.\leo{This could probably be generalized to a set}

For each $M \in \initmach$ and $C = \initcontour(M)$, we define the
{\em closure} of $(C,M)$ as the set of triples of (contour, trace,
trace) as follows:
\[
\begin{array}{rcl}
  \closure{(C,M)} = \{ ( C_i, \seq{N}, \seq{N'})
  & | & C_{i-1} \not = C_i \\
  & , & \seq{(C_i, M_i)} = \induced{C}{M}\\
  & , & \seq{(\_, N_j)} = \induced{C_i}{M_i}\\
  & , & N'_0 \in \variant{C_i}{M_i}\\
  & , & \seq{(\_, N'_j)} = \induced{C_i}{N_0'} \}\\
\end{array}  
\]
\leo{Alternative: We could do this for all $i$, but lose the
  trivial instantiation of noninterference later on}

A trace $\seq{M}$ satisfies the {\em integrity} safety property with
respect to a contour $\contour$, if for all components $\component$
where $\contour(\component) = (\_, \HIGHINT)$, $M_i(K) = M_0(K)$ for
all $i$ in the trace.

A pair of traces $\seq{M}$ and $\seq{N}$ satisfy the {\em
  confidentiality} safety hyperproperty with respect to a contour
$\contour$, if, if the initials states are variants $N_0 \in
\variant{C}{M_0}$, then their observations coincide
$\obstrace(\contour, \seq{M}) = \obstrace(\contour, \seq{N})$.
\leo{Should we cut at minimum length for termination-insensitiveness?}

\leo{From now on I need better terminology, but hopefully you get the
  point} An initial machine state \leo{or contour machine state pair}
is \leo{good} if for every triple $(C, \seq{N}, \seq{N'})$ in
$\closure{(\initcontour(M),M)}$ we have:
\begin{itemize}
\item The traces $\seq{N}$ and $\seq{N'}$ satisfy confidentiality
  with respect to $C$.
\item The traces $\seq{N}$ and $\seq{N'}$ each satisfy integrity
  with respect to $C$.
\end{itemize}

\leo{Combined result}
Noninterference holds for the set of \leo{good} initial machine
states.

\leo{NOTE:
  \begin{itemize}
    \item When the contours don't change this is trivially
      noninterference: only a variation of the initial state is
      added to the trace.
    \item This almost directly coincides with the LSE part
      of the stack-safety property.
  \end{itemize}
}

%A {\em contoured hyperproperty} is a set of triples,
%each consisting of a contour along with a pair of traces.
% 
%\leo{No longer needed:?}
%A {\em contoured lockstep-hyperproperty} is a contoured hyperproperty
%such that each trace in every pair of traces in the set executes the
%same instruction sequence.
% 
%Finally, we can parameterize the traces by some initial application-specific
%state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
%trace comes with more information, for example call sites in stack safety.
%\bcp{If we use this later, we need to say it more formally to be clear about
%  what we mean---e.g., I had thought both the traces and the contour would
%  be functions of $X$!}
